<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Matern correlations}
-->


# Matern correlations

```{r date}
date()
```


```{r packages, results='hide'}
library('geostatsp')
library('gpuR')

```


```{r memoryAvailable, echo=TRUE}
which2.1 = grep('OpenCL 2.1',listContexts()[,'platform'])
if(length(which2.1)) {
  setContext(which2.1[1])
}
gpuInfo()$deviceName
gpuInfo()$maxAllocatableMem/(1024^3)
```

Set up spatial data

```{r setupData}

Nsim = 4
Ngrid = c(10, 10)
Ngrid = c(4,4)
NlocalCache = 1000
Nglobal = c(8L, 8L, 4L)
Nlocal = c(4L, 2L)
theType = "float"

if(gpuInfo()$deviceMemory > 10^10){
   Ngrid = Ngrid * 2
    NlocalCache = 2000
#   Nglobal = c(16L, 16L, 8L)
#   Nlocal = c(8L, 4L)
    theType = "double"
}

myRaster = raster(
    extent(0,Ngrid[1]/Ngrid[2],5,6),
    Ngrid[2], Ngrid[1]
  )
coordsSp = sp::SpatialPoints(xyFromCell(myRaster, 1:ncell(myRaster)))

head(coordsSp@coords)

  coordsGpu = vclMatrix(coordsSp@coords, 
                        nrow(coordsSp@coords), 
                        ncol(coordsSp@coords), type=theType)
```



# Batch matern

```{r paramsBatch}
myParamsBatch = rbind(
  c(shape=1.25, range=0.5, variance = 1.5, nugget = 0, 
             anisoRatio = 1, 
             anisoAngleRadians = 0), 
c(shape=2.15, range=0.25, variance = 2, nugget = 0, 
             anisoRatio = 4, 
             anisoAngleRadians = pi/7),
c(shape=0.55, range=1.5, variance = 2, nugget = 0, 
             anisoRatio = 4, 
             anisoAngleRadians = pi/7),
c(shape=2.15, range=0.5, variance = 2, nugget = 0, 
             anisoRatio = 4, 
             anisoAngleRadians = -pi/7),
c(shape=2.15, range=0.5, variance = 2, nugget = 0, 
             anisoRatio = 2, 
             anisoAngleRadians = pi/4))

myParamsBatch = t(apply(myParamsBatch, 1, geostatsp::fillParam))
myParamsBatch
myParamsBatch = cbind(myParamsBatch, matrix(0, nrow(myParamsBatch), 22-ncol(myParamsBatch)))
paramsGpu =  vclMatrix(myParamsBatch, type=gpuR::typeof(coordsGpu))

outputBatchF = vclMatrix(0, 
    nrow(paramsGpu)*nrow(coordsGpu), nrow(coordsGpu),
    type=gpuR::typeof(paramsGpu))
class(outputBatchF)
```


```{r batchMatern}
  geostatsgpu:::maternBatchBackend(
    outputBatchF,
    coordsGpu, paramsGpu,  
    Nglobal = Nglobal[2:3], 
    Nlocal = Nlocal)
dim(outputBatchF)
```


Compare output to geostatsp's matern

```{r batchMaternCompare}

geostatsp::matern(
    x=coordsSp,
    param=drop(myParamsBatch[1,1:7])
    )[1:4,1:4]
as.matrix(outputBatchF[1:4,1:4])
var4cpu = geostatsp::matern(
    x=coordsSp,
    param=drop(myParamsBatch[4,1:7])
    )
var4gpu = outputBatchF[seq(1+3*nrow(coordsGpu),len=nrow(coordsGpu)), ]
var4cpu[1:4,1:4]
as.matrix(var4gpu)[1:4,1:4]
range(var4cpu - as.matrix(var4gpu))
```

# Cholesky


```{r cholBach}
diagMat = vclMatrix(0, nrow(paramsGpu), ncol(outputBatchF), 
                    type = gpuR::typeof(outputBatchF))
outputChol = deepcopy(outputBatchF)
geostatsgpu:::cholBatchBackend(
  outputChol, diagMat, 
  Nglobal = c(Nglobal[1], Nlocal[2], Nglobal[3]),
  Nlocal = Nlocal,
  NlocalCache = NlocalCache)
```

```{r cholBatchCpu}
cholCpu = as.matrix(chol(var4cpu))
theDiag = diag(cholCpu)
cholCpu = cholCpu %*% diag(1/theDiag)
theDiag = theDiag^2

chol4gpu = as.matrix(outputChol[seq(1+3*nrow(coordsGpu), len=nrow(coordsGpu)),])
chol4gpu[upper.tri(chol4gpu)] = 0
diag(chol4gpu) = 1

chol4gpu[1:4,1:4]
cholCpu[1:4,1:4]
range(cholCpu - as.matrix(chol4gpu))

range(theDiag - as.matrix(diagMat)[4,])
```

# Multiplying lower triangle


create a matrix of independent normals

```{r multLowerSetup}
Nmatrix = nrow(diagMat)
set.seed(0)
zmat = matrix(rnorm(ncol(outputBatchF) * Nsim * Nmatrix), ncol=Nsim)
zmatGpu = vclMatrix(zmat, type = gpuR::typeof(outputChol))
```

compute U = L D^(1/2) Z, where LDLt is the Cholesky of the variance matrix, Z is the independent normals

```{r multLowerGpu}
simMat = deepcopy(-100*zmatGpu)

diagMat = 0*diagMat + 1.0

geostatsgpu:::multiplyLowerDiagonalBatchBackend(
  simMat, outputChol, diagMat, zmatGpu,
  diagIsOne = 1L, # diagonal of L is one
  transformD = "sqrt", # take the square root of each element of D
  Nglobal = Nglobal,
  Nlocal = Nlocal,
  NlocalCache = 200)#NlocalCache)
```

Do the same in R and compare
```{r multLower}
zmatMult = try(diag(as.vector(t(sqrt(as.matrix(diagMat))))) %*% zmat)

rmat = try(mapply(function(DD, xx) {
  xx = xx[seq(1+(DD-1)*ncol(xx),len=ncol(xx)), ]
  xx[upper.tri(xx)] = 0
  diag(xx) = 1
  xx
  },
  DD = 1:Nmatrix,
  MoreArgs = list(xx=as.matrix(outputChol)),
  SIMPLIFY=FALSE))

rmat2 = do.call(Matrix::bdiag, rmat)

multCpu = rmat2 %*% zmatMult

toCompare = as.matrix(simMat) - multCpu
range(toCompare)

toCompare[1:3, ]

badone = which(abs(toCompare[,1]) > 0.01)
length(badone)
if(length(badone)) {
  badone = min(badone)
  print(badone)
  Sbad = badone + seq(-2, 4)
  Sbad = Sbad[Sbad > 0]
  print(toCompare[Sbad, ])
  print(as.matrix(simMat)[Sbad, ])
  print(multCpu[Sbad, ])
}

as.matrix(simMat)[ncol(outputChol) + seq(-2, 4), ]
multCpu[ncol(outputChol) + seq(-2, 4), ]



```


# Plot simulated materns

```{r plotSim, fig.height=4, fig.width=9}
library(raster)
simRaster = brick(myRaster, 
  nl = ncol(simMat)*Nmatrix)
values(simRaster) = as.vector(as.matrix(simMat))

plot(simRaster, nr = Nsim, nc = Nmatrix)

```


# extra info

```{r gpuInfo, echo=TRUE}
gpuR::gpuInfo()
```




# To do

- fix crossprod

    - put crossprod in A and check it's ok
- each workgroup a row

    - use some of diagLocal for within-group sums
- dont store upper triangle in crossprod
- replace * with += for some iterators

