<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Matern correlations}
-->


# Matern correlations

```{r date}
date()
```


```{r packages, results='hide'}
library('geostatsp')
library('gpuR')

Ngrid = c(30, 40)
if(system("hostname", TRUE) == 'englishbreakfast') {
  setContext(2)
#  Ngrid = c(60, 80)
}
```


```{r memoryAvailable, echo=TRUE}
gpuInfo()$deviceName
gpuInfo()$maxAllocatableMem/(1024^3)
```

Set up spatial data

```{r setupData}
seqList = list( 
    x = seq(1,by=1,len=Ngrid[1]), 
    y = seq(201,by=1,len=Ngrid[2])
    )

coordsSp = sp::SpatialPoints(as.matrix(expand.grid(seqList$x, seqList$y)))

head(coordsSp@coords)

myParams = c(shape=2.15, range=2.5, variance = 2, nugget = 0, 
             anisoRatio = 10, 
             anisoAngleRadians = pi/7)
```


```{r simpleTestFloat}

  coordsGpu = vclMatrix(coordsSp@coords, 
                        nrow(coordsSp@coords), 
                        ncol(coordsSp@coords), type="float")
  print(coordsGpu[1:4,1:2])


  outputF = vclMatrix(0, nrow(coordsGpu), nrow(coordsGpu), type='float')

  
system.time(varGpu <- geostatsgpu::maternGpu(
  coordsGpu, output = outputF, param=myParams,
      workgroupSize = c(64L,4L), localSize = c(16L, 4L))
)
```

Use double precision if the GPU supports it

```{r simpleTestDouble}
if(gpuInfo()$double_support) {
  
  coordsGpuD = vclMatrix(coordsSp@coords, 
                         nrow(coordsSp@coords),
                         ncol(coordsSp@coords), type="double")



  outputD = vclMatrix(0, nrow(coordsGpuD), nrow(coordsGpuD), type='double')
  
  system.time(
  varGpuD <- geostatsgpu::maternGpu(
    x= coordsGpuD, output = outputD, param=myParams, 
      workgroupSize = c(64L,4L), localSize = c(16L, 4L))
  )
} else {
  varGpuD = NA
}
```


Create matern matrix on cpu

```{r maternMatGpu}
system.time(varCpu <- geostatsp::matern(coordsSp, myParams))
```

```{r compareCpuGpu}
theProb = c(0.5, 0.9, 0.99, 0.999,1)
quantile(abs(as.matrix(varCpu) - as.matrix(varGpu)), theProb, na.rm=TRUE)
quantile(abs(as.matrix(varCpu) - as.matrix(varGpuD)), theProb, na.rm=TRUE)
endSeq = seq(by=1, len=5, to=nrow(varCpu))
varCpu[endSeq,endSeq]
as.matrix(varGpu[endSeq,endSeq])
as.matrix(varGpuD[endSeq,endSeq])
```

# Batch matern

```{r paramsBatch}
myParamsBatch = rbind(
  c(shape=1.15, range=2.5, variance = 2, nugget = 0, 
             anisoRatio = 1, 
             anisoAngleRadians = 0), 
#myParams, 
myParams * 0.8, myParams*1.2)
myParamsBatch = t(apply(myParamsBatch, 1, geostatsp::fillParam))
myParamsBatch
myParamsBatch = cbind(myParamsBatch, matrix(0, nrow(myParamsBatch), 22-ncol(myParamsBatch)))
paramsGpu =  vclMatrix(myParamsBatch, type='float')


coordsGpu2 = vclMatrix(cbind(1:4, c(1,2,1,2)), 
    type=gpuR::typeof(paramsGpu))
outputBatchF = vclMatrix(0, 
    nrow(paramsGpu)*nrow(coordsGpu2), nrow(coordsGpu2),
    type=gpuR::typeof(paramsGpu))
```


```{r batchMatern}
  geostatsgpu:::maternBatchBackend(
    outputBatchF,
    coordsGpu2, paramsGpu,  
    Nglobal = c(4L,4L), 
    Nlocal = c(2L, 2L))
```


Compare output to geostatsp's matern
```{r batchMaternCompare}
geostatsp::matern(
    x=SpatialPoints(as.matrix(coordsGpu2)),
    param=drop(myParamsBatch[1,1:6])
    )[1:4,1:4]
as.matrix(outputBatchF[1:4,1:4])
geostatsp::matern(
    x=SpatialPoints(as.matrix(coordsGpu2)),
    param=drop(myParamsBatch[2,1:6])
    )[1:4,1:4]
as.matrix(outputBatchF[nrow(coordsGpu2)+1:4,1:4])
```


```{r cholBach}
cholCpu = t(chol(as.matrix(
  outputBatchF[1:ncol(outputBatchF), ]
  )))
theDiag = diag(cholCpu)
cholCpu = cholCpu %*% diag(1/theDiag)
theDiag = theDiag^2


diagMat = vclMatrix(0, nrow(paramsGpu), ncol(outputBatchF), type = gpuR::typeof(outputBatchF))
geostatsgpu:::cholBatchBackend(
  outputBatchF, diagMat, 
   c(4L,4L, 2L),
   c(2L,2L, 1L))
as.matrix(outputBatchF)
cholCpu
as.matrix(diagMat)
theDiag
```

# Cholesky


```{r cholGeostatsGpu, eval=FALSE}

varGpuD2 = deepcopy(varGpuD)
varGpuDiag = vclVector(0, nrow(varGpuD), type="double")


	file = "/home/patrick/research/diseasemapping/pkg/geostatsgpu/inst/CL/cholGpu.cl"
	if(!file.exists(file)) {
  	file <- system.file("CL", "cholGpu.cl", package = "geostatsgpu")
	}
	kernel <- readChar(file, file.info(file)$size)

	diagWorking = deepcopy(varGpuDiag) 
	diagTimesRow = deepcopy(varGpuDiag) 
	
system.time(geostatsgpu:::cpp_cholGpu(
  xR = varGpuD2,
  DR = varGpuDiag, 
  diagWorkingR = diagWorking, 
  diagTimesRowOfAR = diagTimesRow, 
  MCglobal = 256, 
  MClocal = 8, 
  localStorage = floor(0.9*gpuR::gpuInfo()$localMem/8), 
  colGroupwise = nrow(varGpuD)-64,
  Ncrossprod = 32, 
  verbose = FALSE, 
  kernelR = kernel))

system.time(theChol <- chol(as.matrix(varGpuD)) )

quantile(as.vector(varGpuDiag) - diag(theChol)^2)

theChol = Matrix::Diagonal(nrow(theChol), 1/diag(theChol) ) %*% theChol
as.matrix(varGpuD2[1:5,1:5])
theDiff = theChol - as.matrix(varGpuD2)
theDiff[lower.tri(theDiff)] = 0
quantile(theDiff)


```

Using ViennaCL's LU decomposition
```{r cholViennaclLu, eval=FALSE}
varMat3 = gpuR::deepcopy(varGpuD)
system.time(varCholVcl <- geostatsgpu::luGpu(varMat3))

system.time(theChol <- chol(as.matrix(varGpuD)) )

quantile(as.vector(varCholVcl$D) - diag(theChol)^2)
```


# Integrated Cholesky


```{r maternMatGpuC, eval=FALSE}
system.time(varCpuC <- geostatsp::matern(coordsSp, myParams, type='cholesky'))
system.time(varGpuC <- geostatsgpu::maternGpu(x=vclMatrix(coordsSp@coords, type='float'), param=myParams, 
    form='cholesky'))
system.time(varGpuCD <- geostatsgpu::maternGpu(x=vclMatrix(coordsSp@coords, type='double'), param=myParams, 
    form='cholesky'))
quantile(varCpuC - as.matrix(varGpuC), na.rm=TRUE)
quantile(varCpuC - as.matrix(varGpuCD), na.rm=TRUE)
```


# Cholesky again

Cholesky on the CPU
```{r cholCpu, eval=FALSE}
system.time(varCholGpu <- chol(varCpu))
```

Using `geostatsgpu`'s Cholesky
```{r cholGpu, eval=FALSE}
varMat1 = gpuR::deepcopy(varGpuD)
try(system.time(varCholGpu <- geostatsgpu:::cholGpu(varMat1)))
```

Using `gpuR`'s Cholesky
```{r cholGpurPackage, eval=FALSE}
varMat2 = gpuR::deepcopy(varGpuD)
system.time(varCholGpuR <- gpuR:::chol.vclMatrix(varMat2))
```


```{r cholGpuCompare, eval=FALSE}
xchold = diag(varGpuD)
cholTest = t((diag(1/xchold) %*% varCholGpu))
DfromR = xchold^2
quantile(DfromR - as.vector(varCholGpu$D), na.rm=TRUE)
```



# extra info

```{r gpuInfo, echo=TRUE}
gpuR::gpuInfo()
```




# To do

- fix crossprod

    - put crossprod in A and check it's ok
- each workgroup a row

    - use some of diagLocal for within-group sums
- dont store upper triangle in crossprod
- replace * with += for some iterators

