<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Matern correlations}
-->


# Matern correlations

```{r date}
date()
```



```{r packages, results='hide', message='hide'}
library('gpuR')
```


```{r memoryAvailable, echo=TRUE}
listContexts()[,'platform']
which2.1 = grep('OpenCL 2.1',listContexts()[,'platform'])
if(length(which2.1)) {
  setContext(which2.1[1])
}
gpuInfo()$deviceName
gpuInfo()$maxAllocatableMem/(1024^3)
```

Set up spatial data

```{r setupData}

   Nsim = 4
   Ngrid = c(4,5)
   NlocalCache = 300
    NglobalMatern = c(64,8) 
    NlocalMatern = c(8, 4)
   NglobalChol =  c(16,8,4)
    NglobalChol =  c(64,8)
    NlocalChol = c(16, 8)
    NglobalRandom = c(64,4)
  NglobalMult = c(16,4,4)   
  NlocalMult = c(4,4,1)   
theType = "float"


NlocalMatern = c(2,2)


if(gpuR::gpuInfo()$deviceMemory > 10^10){
   Ngrid = c(4,5)
   Nsim = 6
   NlocalCache = 4000
   NglobalMatern = c(256,8) 
   NlocalMatern = c(32, 8)
   NglobalChol =  c(128,8)
   NlocalChol = c(16, 8)
   NglobalRandom = c(512,8)
   NglobalMult = c(64,8,8)   
   NlocalMult = c(8,8,1)
   theType = "double"
}

myRaster = raster::raster(
    raster::extent(0,Ngrid[1]/Ngrid[2],5,6),
    Ngrid[1], Ngrid[2]
  )
coordsSp = sp::SpatialPoints(raster::xyFromCell(myRaster, 1:raster::ncell(myRaster)))
head(coordsSp@coords,15)

```



# Batch matern

```{r paramsBatch}
myParamsBatch = rbind(
c(shape=5.9, range=1, variance = 1, nugget = 0, anisoRatio = 1, anisoAngleRadians = 0), 
c(shape=6, range=1, variance = 1, nugget = 0, anisoRatio = 1, anisoAngleRadians = 0)) 
# c(shape=3, range=5, variance = 1, nugget = 0, anisoRatio = 1, anisoAngleRadians = 0), 
# c(shape=4, range=5, variance = 1, nugget = 0, anisoRatio = 1, anisoAngleRadians = 0), 
# c(shape=5, range=5, variance = 1, nugget = 0, anisoRatio = 1, anisoAngleRadians = 0) )

myParamsBatch = t(apply(myParamsBatch, 1, geostatsp::fillParam))
myParamsBatch
myParamsBatch = cbind(myParamsBatch, matrix(0, nrow(myParamsBatch), 22-ncol(myParamsBatch)))
```

```{r outputBatchCreate, eval=T}
library(gpuR)
coordsGpu = vclMatrix(coordsSp@coords, nrow(coordsSp@coords), ncol(coordsSp@coords), type=theType)
paramsGpu = vclMatrix(myParamsBatch, type=gpuR::typeof(coordsGpu))
outputBatchF = vclMatrix(0, nrow(paramsGpu)*nrow(coordsGpu), nrow(coordsGpu), type=gpuR::typeof(paramsGpu))

class(outputBatchF)
```

```{r batchMatern, eval=T}
system.time(gpuRandom:::maternBatchBackend(
    outputBatchF,
    coordsGpu, 
    paramsGpu,  
    Nglobal = NglobalMatern, 
    Nlocal = NlocalMatern))
dim(outputBatchF)

gpuRandom:::maternBatchBackend(
    outputBatchF,
    coordsGpu, 
    paramsGpu,  
    Nglobal = NglobalMatern, 
    Nlocal = NlocalMatern)
fromCpu = geostatsp::matern(
    x=coordsSp,
    param=drop(myParamsBatch[1,1:7])
    )
fromCpu[1:8,1:8]
as.matrix(outputBatchF[1:8,1:8])
#as.matrix(1000*outputBatchF[1:8,1:8])
as.matrix(outputBatchF[21:28,1:8])

#################
#upper
thisx = as.matrix(dist(coordsSp@coords)/5)
#lower
ln_half_x = log(thisx/2)

####compare
as.matrix(outputBatchF[1:8,1:8])
thisx[1:8,1:8]
as.matrix(outputBatchF[1:8,1:8])
ln_half_x[1:8,1:8]

#lower
lower<-outer(coordsSp@coords[,1], coordsSp@coords[,1], FUN = '-')
lower[1:8,1:8]
```









```{r batchMaternCompare, include=TRUE, eval=TRUE}
# Compare output to geostatsp's matern

fromCpu = geostatsp::matern(
    x=coordsSp,
    param=drop(myParamsBatch[1,1:7])
    )
fromCpu[1:4,1:4]
as.matrix(outputBatchF[1:4,1:4])

quantile(fromCpu - as.matrix(outputBatchF[1:ncol(outputBatchF),]))


var4cpu = geostatsp::matern(
    x=coordsSp,
    param=drop(myParamsBatch[4,1:7])
    )

var4gpu = as.matrix(outputBatchF[seq(1+3*nrow(coordsGpu),len=nrow(coordsGpu)), ])
var4cpu[1:4,1:4]
var4gpu[1:4,1:4]
quantile(var4cpu - var4gpu)
```

# Cholesky
```{r cholBach, eval = F}
diagMat = vclMatrix(0, nrow(paramsGpu), ncol(outputBatchF), type = gpuR::typeof(outputBatchF))

outputChol = vclMatrix(0, nrow(outputBatchF), ncol(outputBatchF), type = gpuR::typeof(outputBatchF))

system.time(gpuRandom::cholBatch(
  outputChol, diagMat, numbatchD=nrow(myParamsBatch),
  Nglobal = NglobalChol,
  Nlocal = NlocalChol,
  NlocalCache = NlocalCache))

as.matrix(outputChol[1:12,1:7])
as.matrix(diagMat[1:12,1:7])
```

```{r cholBatchCpu, include=FALSE, eval=F}
cholCpu = as.matrix(chol(var4cpu))
theDiag = diag(cholCpu)
cholCpu = cholCpu %*% diag(1/theDiag)
theDiag = theDiag^2

chol4gpu = as.matrix(outputChol[seq(1+3*nrow(coordsGpu), len=nrow(coordsGpu)),])
chol4gpu[upper.tri(chol4gpu)] = 0
diag(chol4gpu) = 1

chol4gpu[1:4,1:4]
cholCpu[1:4,1:4]
range(cholCpu - as.matrix(chol4gpu))

range(theDiag - as.matrix(diagMat)[4,])
```

# Multiplying lower triangle


create a matrix of independent normals


```{r randomNormalsSim, eval=F}
set.seed(1)
streams = gpuRandom::CreateStreams(prod(NglobalRandom))
streamsGpu = vclMatrix(streams, type='integer')

zMatGpu = vclMatrix(nrow=as.integer(ncol(outputBatchF) * nrow(myParamsBatch)),
                    ncol=as.integer(Nsim),
                    type = gpuR::typeof(outputBatchF))
dim(zMatGpu)
gpuRandom:::gpuRnBackend(zMatGpu, streamsGpu, NglobalRandom, "normal")
#system.time(gpuRandom:::gpuRnBackend(zMatGpu, streamsGpu, NglobalRandom, "normal"))

```

```{r checkRange, eval=F, include=FALSE}
streams[1:3,1:6]
as.matrix(streamsGpu[1:3,1:6])
zMatCpu = as.matrix(zMatGpu)
hist(zMatCpu)
qqnorm(as.vector(zMatCpu))
```




compute U = L D^(1/2) Z, where LDLt is the Cholesky of the variance matrix, Z is the independent normals

```{r multLowerGpu, eval = F}
simMat = deepcopy(-100*zMatGpu)

system.time(
gpuRandom:::multiplyLowerDiagonalBatchBackend(
  simMat, outputChol, diagMat, zMatGpu,
  diagIsOne = TRUE, # diagonal of L is one
  transformD = "sqrt", # take the square root of each element of D
  Nglobal = NglobalMult,
  Nlocal = NlocalMult,
  NlocalCache = NlocalCache)
)
```



# Plot simulated materns

```{r plotSimNames,eval=F }
theNames = apply(expand.grid('par',1:nrow(paramsGpu), 'sim', 1:Nsim), 1, paste, collapse='')
```

```{r plotSimSetup, eval = F}
library(raster)
simRaster = brick(myRaster, 
  nl = ncol(simMat)*nrow(paramsGpu))
system.time(values(simRaster) <- as.vector(as.matrix(simMat)))
names(simRaster) = theNames
```

```{r plotSim, fig.height=5, fig.width=4, out.width="paste0(98/nrow(paramsGpu), '%')", fig.cap='simulations', fig.subcap = 'theNames', eval = F}
par(mar=rep(0.1, 4))
for(D in names(simRaster)) {
  plot(extent(simRaster))
  plot(simRaster[[D]], legend=FALSE, add=TRUE)
}
  
```



outputChol a = cbind(Y, 1), solve for a

# Backsolve

```{r backsolve, echo=TRUE, eval=F}
  
#solveMat = -99 +0*simMat
solveMat=vclMatrix(0, nrow(simMat),ncol(simMat),type=gpuR::typeof(outputChol))

gpuRandom:::backsolveBatchBackend(
  solveMat, outputChol, simMat, 
  diagIsOne = TRUE,   
  Nglobal = NglobalChol,
  Nlocal = NlocalChol,
  NlocalCache = NlocalCache)

NglobalChol
NlocalChol
NglobalChol/NlocalChol

as.matrix(simMat[1:12,])

as.matrix(solveMat[1:12,])

L = as.matrix(outputChol[1:ncol(outputChol), ])
L[!lower.tri(L, diag=TRUE)] = 0
diag(L) = 1
L[1:4,1:4]

solve(L, as.matrix(simMat[1:ncol(outputChol),]))[1:12, ]

```


# matern again

```{r matern test Nov23,2020, eval=F}
library('geostatsp')
data('swissRain')
sr2 = swissRain[1:20,]

# extract variables for modeling
sr2$sqrtY = sqrt(sr2$rain)
sr2$elev = raster::extract(swissAltitude, sr2)


swissFitIso = likfitLgm(
  data=sr2,
  formula=sqrtY ~ elev,
  param=c(range=15,shape=1,nugget=0),
  reml=FALSE,
  verbose=FALSE
)

swissFitIso$parameters

# newParamList = list(
#   range=seq(15, 50 , len=3)*1000,
#   nugget = seq(0.01,0.5,len=3)
# )
# newParam= do.call(expand.grid, newParamList)
# 
# otherparams = c(shape=1, variance = 1.5, anisoRatio = 1, anisoAngleRadians = 0)
# otherparams = matrix(rep(otherparams,each=nrow(newParam)),nrow=nrow(newParam))
# colnames(otherparams) <- c("shape","variance", "anisoRatio", "anisoAngleRadians")
# paramsBatch0 = cbind(newParam, otherparams)
paramsBatch0 = rbind(c(shape=1.25, range=5, variance = 1, nugget = 0, anisoRatio = 1, anisoAngleRadians = 0),
                     c(shape=1.25, range=5, variance = 1, nugget = 0, anisoRatio = 1, anisoAngleRadians = 0),
                     c(shape=1.25, range=5, variance = 1, nugget = 0, anisoRatio = 1, anisoAngleRadians = 0),
                     c(shape=1.25, range=5, variance = 1, nugget = 0, anisoRatio = 1, anisoAngleRadians = 0))

paramsBatch0 = t(apply(paramsBatch0, 1, geostatsp::fillParam))
paramsBatch = vclMatrix(cbind(paramsBatch0, matrix(0, nrow(paramsBatch0), 22-ncol(paramsBatch0))),type="double")
paramscpu<-as.matrix(paramsBatch)
workgroupSize=c(64,16,16)
localSize = c(4,4,1)
NlocalCache=50

#betas=NULL
#form="ml"
type="double"
#covariates = model.matrix(swissFitIso$model$formula, data=swissFitIso$data)
temp = model.frame(swissFitIso$model$formula, data=swissFitIso$data)
response=temp[,as.character(attributes(terms(temp))$variables)[2]]
#yX=vclMatrix(cbind(response,covariates),type=type)
coordsGpu<-vclMatrix(sr2@coords/1000,type=type)
n = length(response)
#p = ncol(covariates)


 # form = c(loglik=1, ml=2, mlFixSigma=3, mlFixBeta=4, reml=5, remlPro=6)[form]
 # rowbatch = nrow(paramsBatch)
 # colbatch = 1     #colbatch = ncol(y)
   
 # localSizechol<-localSize
 # localSizechol[2]<-workgroupSize[2]
    
 Vbatch = vclMatrix(0, nrow(paramsBatch)*n, n, type = type)
 diagMat = vclMatrix(0, nrow(paramsBatch), n, type = type) 
     
     #ycpu<-as.matrix(y)
     #xcpu<-as.matrix(X)
     
  # 
 gpuRandom:::maternBatchBackend(Vbatch, coordsGpu, paramsBatch,  workgroupSize, localSize)
 vbatchcpu<-as.matrix(Vbatch)
 vbatchcpu[1:6,1:6]
 var4cpu[1:6,1:6]
 diag(vbatchcpu[1:20,])
     
 
 var4cpu = geostatsp::matern(
      x=sr2,
      param=drop(paramsBatch0[1,1:7]))
 
 quantile(var4cpu-vbatchcpu[1:20,])
     
 dim(var4cpu)
 vbatchcpu[1:20,]-var4cpu
 diag(vbatchcpu[1:20,])-diag(var4cpu)


#################
#upper
thisx = as.matrix(dist(sr2@coords/1000)/5)

#lower
ln_half_x = log(thisx/2)

vbatchcpu[1:7,1:7]
thisx[1:7,1:7]
vbatchcpu[1:7,1:7]
ln_half_x[1:7,1:7]

if anisotriopc then rotate coordinates before takking distance matrix.

see line 283 in maternBatch.cpp 





    


```

problem is maternShort call

1- check order of parameters
2- check the fill22params
3- comment out bits of kernel and retun distance or other things

```{r batchMaternSwussRaub, eval=T}
data('swissRain', package='geostatsp')
sr2 = swissRain[1:4,]
sr2@coords = sr2@coords
dist(sr2@coords)
myParamsSR = geostatsp::fillParam(c(
  variance=2,nugget=1, anisoAngleRadians=0,
  range=12*1000, shape=1, anisoRatio = 1))

myParamsBatchSR1 = rbind(myParamsSR, myParamsSR)
myParamsBatchSR = cbind(myParamsBatchSR1, 
        matrix(0, nrow(myParamsBatchSR1), 22-ncol(myParamsBatchSR1)))

outputSwissRain = vclMatrix(0, 
        nrow(myParamsBatchSR)*length(sr2),length(sr2),type='double')
gpuRandom:::maternBatchBackend(
    var=outputSwissRain,
    coords=vclMatrix(sr2@coords, type='double'), 
    param=vclMatrix(myParamsBatchSR, type='double'),  
    Nglobal = c(1,1,1), 
    Nlocal = c(1,1,1))
geostatsp::matern(sr2, myParamsSR)
as.matrix(outputSwissRain)
```
# extra info

```{r gpuInfo, echo=TRUE}
gpuR::gpuInfo()
```





